---
title: "Hands-on Exercise 2"
subtitle: "Beyond ggplot2 Fundamentals"
date: January 15, 2024
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  toc: true
---

## Getting Started

### Loading R packages

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

### Importing data

```{r}
exam_data <- read_csv("data/exam_data.csv")
```

## Beyong ggplot2 Annotation: ggrepel

[ggrepel](https://ggrepel.slowkow.com/) provides geoms for [ggplot2](https://ggplot2.tidyverse.org/) to repel overlapping text labels:

-   [`geom_text_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel)
-   [`geom_label_repel()`](https://ggrepel.slowkow.com/reference/geom_text_repel)

`geom_text_repel` adds text directly to the plot. `geom_label_repel` draws a rectangle underneath the text, making it easier to read. The text labels repel away from each other and away from the data points.

```{r}
ggplot(data= exam_data,
       aes(x= MATHS,
           y= ENGLISH,
           colour= GENDER)) +
  geom_point() +
  geom_label_repel(aes(label=ID),
                   label.padding = 0.2,
                   label.size = 0.2) +
  xlim(0,100) +
  ylim (0,100)

```

## Beyong ggplot2 Themes

[gglot2](https://ggplot2.tidyverse.org/) comes with 10 [built-in themes](https://ggplot2.tidyverse.org/reference/#themes). Theme could be modified by using [`theme()`](https://ggplot2.tidyverse.org/reference/theme.html) and [`element_`](https://ggplot2.tidyverse.org/reference/element.html).

The code chunk below plots basic plot with modification to color and fill of the bar before any themes were applied.

```{r}
ggplot(data= exam_data,
       aes(x= SCIENCE))+
  geom_histogram(binwidth = 5,
                 boundary = 5,
                 color= "black",
                 fill= "white") 
```

### Working with ggtheme package

[`ggthemes`](https://cran.r-project.org/web/packages/ggthemes/) provides extra theme, geoms, and scales for ggplot2. Refer to these (i) [link](https://jrnold.github.io/ggthemes/reference/index.html) (ii) [link](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/) for examples.

```{r}
ggplot(data= exam_data,
       aes(x= SCIENCE))+
  geom_histogram(binwidth = 5,
                 boundary = 5,
                 color= "black",
                 fill= "white") +
  theme_wsj(base_size = 10,
             color = "brown",
             base_family = "sans",
             title_family = "mono") +
  ggtitle("Distribution of Maths scores") +
  theme(plot.title= element_text(size= 15))
```

### Working with hrbrthemes package

[hrbrthemes](https://cinc.rud.is/web/packages/hrbrthemes/) is a very focused package that provides typography-centric themes and theme components for ggplot2. Consult this [vignette](https://cran.r-project.org/web/packages/hrbrthemes/vignettes/why_hrbrthemes.html) to learn more.

```{r}
ggplot(data= exam_data,
       aes(x= SCIENCE))+
  geom_histogram(binwidth = 5,
                 boundary = 5,
                 color = "black",
                 fill = "bisque3",
                 alpha = 0.6,
                 linewidth =0.3) +
  ggtitle("Distribution of Science scores") +
  theme_ipsum(axis_title_size = 12,
              base_size = 10,
              grid_col = "grey80",
              grid = "Y")
```

::: {.callout-tip appearance="simple"}
## What can we learn from the code chunk above?

-   `axis_title_size` argument is used to increase the font size of the axis title to 18,
-   `base_size` argument is used to increase the default axis label to 15, and
-   `grid` argument is used to remove the x-axis grid lines.
:::

## Beyong Single Graph

[Patchwork](https://patchwork.data-imaginist.com/) which is specially designed for combining separate ggplot2 graphs into a single figure.

Patchwork package has a very simple syntax where we can create layouts super easily. Here’s the general syntax that combines:

-   Two-Column Layout using the Plus Sign +.
-   Parenthesis () to create a subplot group.
-   Two-Row Layout using the Division Sign /

To learn more about, refer to [`Plot Assembly`](https://patchwork.data-imaginist.com/articles/guides/assembly.html).

### Combining 2 graphs using ' + '

```{r}
p1 <- ggplot(data= exam_data,
       aes(x= MATHS))+
  geom_histogram(binwidth = 5,
                 boundary = 5,
                 color = "black",
                 fill = "bisque",
                 alpha = 0.6,
                 linewidth =0.3) +
  labs(x= "Maths score",
       y= "No. of students",
       title="Distribution of Maths scores") +
    theme(plot.title= element_text(size= 12, hjust= 0.5),
          axis.title= element_text(size= 10),
          panel.background = element_rect(fill= "grey90"))

p2 <- ggplot(data= exam_data,
       aes(x= SCIENCE))+
  geom_histogram(binwidth = 5,
                 boundary = 5,
                 color = "black",
                 fill = "bisque4",
                 alpha = 0.6,
                 linewidth =0.3) +
  labs(x= "Science score",
       y= "No. of students",
       title="Distribution of Science scores") +
    theme(plot.title= element_text(size= 12, hjust= 0.5),
          axis.title= element_text(size= 10),
          panel.background = element_rect(fill= "grey90"),
          axis.title.y = element_blank())
          

p1 + p2

```

### Combining 3 graphs using ' \| ' ' / ' ' ( ) '

```{r}
p3 <- ggplot(data = exam_data,
             aes(MATHS, SCIENCE)) +
  geom_point(size = 2,
             color = "khaki4",
             alpha = 0.8) +
  geom_smooth(method = lm,
              size = 0.8,
              color = "grey20") +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100)) +
  labs(title = "Correlation Plot",
       subtitle = "Science vs Maths scores for Primary 3") +
  theme(plot.title=element_text(size= 12),
        plot.subtitle=element_text(size= 10),
        axis.title = element_text(size= 10))

(p1 / p2) | p3
```

### Adding [annotation](https://patchwork.data-imaginist.com/articles/guides/annotation.html)

```{r}
((p1 / p2) | p3) + 
  plot_annotation(title= "Distribution and correlation of Maths and Science scores for Primary 3 students",
                  tag_levels = 'i')
```

### Insert Figures with [`Inset_element`](https://patchwork.data-imaginist.com/reference/inset_element.html)

```{r}
p3 + inset_element(p2, 
                   left = 0.02, 
                   bottom = 0.5, 
                   right = 0.4, 
                   top = 1)
```

### Creating a composite figure by using patchwork and ggtheme

```{r}
patchwork <- (p1 / p2) | p3
patchwork & theme_stata()
```
