---
title: "Take-home Exercise 1"
subtitle: "Creating data visualisation beyond default"
date: January 17, 2024
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  toc: true
format: 
  html:
    code-fold: false
editor: visual
---

## Overview

OECD education director Andreas Schleicher shared in a BBC article that “Singapore managed to achieve excellence without wide differences between children from wealthy and disadvantaged families.” (2016) However, there are public concern about the inequality, particularly in the context between elite schools and neighborhood school, between families with higher socioeconomic status and those with relatively lower socioeconomic status and immigration and non-immigration families.

### Objectives

This analysis using the 2022 PISA data aims to explore these aspects in Singapore's context. The objective of this study is to adopt appropriate Exploratory Data Analysis (EDA) methods and ggplot2 functions to reveal:

-   the distribution of Singapore students’ performance in mathematics, reading, and science, and
-   the relationship between these performances with schools, gender and socioeconomic status of the students.

## Data preparation

### Loading R packages

First, following code chunk with `pacman::p_load()` function is used to load required R packages into our working environment:

```{r}
#| code-fold: false
pacman::p_load(tidyverse, haven, patchwork, ggdist, ggrain, ggridges)
```

### Importing PISA data

The code chunk below uses [`read_sas()`](https://haven.tidyverse.org/reference/read_sas.html) of [haven](https://haven.tidyverse.org/) to import PISA data into working environment

```{r}
#| eval: false
#| code-fold: false
stu_qqq <- read_sas("data/cy08msp_stu_qqq.sas7bdat")
```

The data table consists of 613,744 observations and 1,279 variables.

### Data extracting

Referring to PISA code book, it is noted that the data table include results from 81 participating countries. As this study focus on Singapore's context, relevant observations is filtered using `filter()` on the country code variable "CNT".

```{r}
#| eval: false
#| code-fold: false
stu_qqq_SG <- stu_qqq %>%
  filter(CNT == "SGP")
```

The filtered data is then saved in R Data Format and re-imported into the working environment.

```{r}
#| eval: false
#| code-fold: false
write_rds(stu_qqq_SG, "data/stu_qqq_SG.rds")
```

```{r}
#| code-fold: false
stu_qqq_SG <- read_rds("data/stu_qqq_SG.rds")
stu_qqq_SG
```

After completing the previous steps, we are left with a substantial data table consisting 1,279 variables. At this point, it is essential to narrow down these variables to focus on the most relevant ones. By examining the code book, the following variables were chosen to align with the objective of this study:

+---------------+-----------------------------------------------+------------------------------------------------+
| Variable name | Label                                         | Description                                    |
+===============+===============================================+================================================+
| CNTSTUID      | Intl. Student ID                              | Student ID of participating students           |
+---------------+-----------------------------------------------+------------------------------------------------+
| STRATUM       | Stratum ID 5-character\                       | Type of School\                                |
|               | (cnt + original stratum ID)                   | SGP01 : Public / Secondary\                    |
|               |                                               | SGP03 : Private / Secondary                    |
+---------------+-----------------------------------------------+------------------------------------------------+
| ST004D01T     | Student (Standardized) Gender                 | Gender\                                        |
|               |                                               | 1 : Female\                                    |
|               |                                               | 2 : Male                                       |
+---------------+-----------------------------------------------+------------------------------------------------+
| ESCS          | Index of economic, social and cultural status | Measurement of student's socio-economic status |
+---------------+-----------------------------------------------+------------------------------------------------+
| PV1MATH       | Plausible Value 1 in Mathematics              | Student's performance for Mathematics          |
+---------------+-----------------------------------------------+------------------------------------------------+
| PV1READ       | Plausible Value 1 in Reading                  | Student's performance for Reading              |
+---------------+-----------------------------------------------+------------------------------------------------+
| PV1SCIE       | Plausible Value 1 in Science                  | Student's performance for Science              |
+---------------+-----------------------------------------------+------------------------------------------------+

We use `select()` and `rename()`from [dplyr](https://dplyr.tidyverse.org/articles/dplyr.html) to select the column and rename the variable for clarity.

```{r}
stu_qqq_SG_selected <- stu_qqq_SG %>%
  select('CNTSTUID',
         'STRATUM',
         'ST004D01T',
         'ESCS',
         'PV1MATH',
         'PV1READ',
         'PV1SCIE') %>%
  rename(StudentID = CNTSTUID,
         TypeofSchool = STRATUM,
         Gender = ST004D01T,
         MATH = PV1MATH,
         READ = PV1READ,
         SCIENCE = PV1SCIE)
```

### Summary Statistics of data table

::: panel-tabset
#### Data table

```{r}
stu_qqq_SG_selected
```

#### Summary

```{r}
summary(stu_qqq_SG_selected)
```

#### Duplicate

Check for duplicated value

```{r}
stu_qqq_SG_selected[duplicated(stu_qqq_SG_selected),]
```

#### Missing

Checking for missing data

```{r}
stu_qqq_SG_selected %>% summarise_all(funs(sum(is.na(.))))
```

The output shows 47 missing values for ESCS, the observations were dropped using `drop_na()`:

```{r}
stu_qqq_SG_selected2 <- stu_qqq_SG_selected %>% drop_na()
```
:::

> ⚠️ **Data quality issues**

-   **Inappropriate column types**\
    Following columns were converted to appropriate type:
    -   "StudentID" : convert from `<dbl>` to `<chr>`
    -   "TypeofSchool" : convert from `<chr>` to `<fct>`
    -   "Gender" : convert from `<dbl>` to `<fct>`
-   **Non-Descriptive Values**\
    Values from "Gender" and "TypeofSchool" were recoded to descriptive values using `fct_recode` to improve clarity.
-   **Disaggregated ESCS Index**\
    Binning of the ESCS index into categories of most disadvantaged to the most advantaged groups can facilitate more granular analysis and provide meaningful insights. 4 groups of students of equal size were created using `cut_number()`, each group comprises 25% of the population of 15-year-old students in Singapore.

```{r}
#| code-fold: true
stu_qqq_SG_converted <- stu_qqq_SG_selected2 %>%
  
  # change column type
  mutate(StudentID = as.character(StudentID),         
         TypeofSchool = as.factor(TypeofSchool),
         Gender = as.factor(Gender)) %>%
  # recode non-descriptive values  
  mutate(Gender = fct_recode (Gender,
                              "Female" = "1",
                              "Male" = "2"),
           TypeofSchool = fct_recode (TypeofSchool,
                              "Public" = "SGP01",
                              "Private" = "SGP03"),
  # binning of disaggregated data       
         binned_ESCS = cut_number(stu_qqq_SG_selected2$ESCS, 
                                    n = 4, 
                                    labels = c("Disadvantaged",
                                               "Slightly Disadvantaged",
                                               "Slightly Advantaged",
                                               "Advantaged")))
```

Following final data table is used for Exploratory data analysis:

```{r}
stu_qqq_SG_converted
```

## Exploratory Data Analysis

### Distribution of Performance in Mathematics, Reading and Science

```{r}
#| code-fold: true

# Distribution of Performance in Mathematics
P1 <- ggplot(data = stu_qqq_SG_converted,
       aes(x = MATH)) +
  geom_density(color = "#459395", size = 0.6, fill= "#459395", alpha = 0.4) +
  coord_cartesian(xlim = c(0,1000)) +
  geom_vline(aes(xintercept = mean(MATH)),
             color = "red", alpha = 0.8, linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = 400, y = 0.0035,
           label = paste("Mean=", 
                         round(mean(stu_qqq_SG_converted$MATH, na.rm=T), 2)),
           color = "red", size = 3) +
  geom_vline(aes(xintercept = median(MATH)),
             color= "grey50", linewidth = 0.7, linetype = "solid") +
  annotate("text", x = 800, y = 0.0035,
           label = paste("Median=", 
                         round(median(stu_qqq_SG_converted$MATH, na.rm=T), 2)),
           color = "grey20", size = 3) +  
  geom_boxplot(width = 0.0005, fill = "white", alpha = 0.5,
               position = position_nudge(y = -0.0005)) +
  theme_minimal()+
  labs(title="Distribution of Performance in Mathematics") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12),
        axis.text = element_text(size= 8)) 

# Distribution of Performance in Reading
P2 <- ggplot(data = stu_qqq_SG_converted,
       aes(x = READ)) +
  geom_density(color = "#EB7C69", size = 0.6, fill= "#EB7C69", alpha = 0.4) +
  coord_cartesian(xlim = c(0,1000)) +
  geom_vline(aes(xintercept = mean(READ)),
             color = "red", alpha = 0.8, linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = 400, y = 0.0035,
           label = paste("Mean=", 
                         round(mean(stu_qqq_SG_converted$READ, na.rm=T), 2)),
           color = "red", size = 3) +
  geom_vline(aes(xintercept = median(READ)),
             color= "grey50", linewidth = 0.7, linetype = "solid") +
  annotate("text", x = 800, y = 0.0035,
           label = paste("Median=", 
                         round(median(stu_qqq_SG_converted$READ, na.rm=T), 2)),
           color = "grey20", size = 3) +  
  geom_boxplot(width = 0.0005, fill = "white", alpha = 0.5,
               position = position_nudge(y = -0.0005)) +
  theme_minimal()+
  labs(title="Distribution of Performance in Reading",
       y = "density") +
  theme(axis.title.x = element_blank(),
        plot.title=element_text(size= 12),
        axis.text = element_text(size= 8)) 
  
# Distribution of Performance in Science
P3 <- ggplot(data = stu_qqq_SG_converted,
       aes(x = SCIENCE)) +
  geom_density(color = "#FDA638", size = 0.6, fill= "#FDA638", alpha = 0.4) +
  coord_cartesian(xlim = c(0,1000)) +
  geom_vline(aes(xintercept = mean(SCIENCE)),
             color = "red", alpha = 0.8, linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = 400, y = 0.0035,
           label = paste("Mean=", 
                         round(mean(stu_qqq_SG_converted$SCIENCE, na.rm=T), 2)),
           color = "red", size = 3) +
  geom_vline(aes(xintercept = median(SCIENCE)),
             color= "grey50", linewidth = 0.7, linetype = "solid") +
  annotate("text", x = 800, y = 0.0035,
           label = paste("Median=", 
                         round(median(stu_qqq_SG_converted$SCIENCE, na.rm=T), 2)),
           color = "grey20", size = 3) +  
  geom_boxplot(width = 0.0005, fill = "white", alpha = 0.5,
               position = position_nudge(y = -0.0005)) +
  theme_minimal()+
  labs(title="Distribution of Performance in Science") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12),
        axis.text = element_text(size= 8)) 


P1 / P2 / P3
```

The density plot for performance in Mathematics, reading and science shows a predominantly normal distribution with a slight left skew, as indicated by mean scores being lower than medians in all subjects. The box plot below reveals tight interquartile range in all subjects, which indicates that 50% of the students performed within a relatively narrow range, emphasizing a strong central tendency.

Across three subjects, mathematics showed the highest performance on average, followed by Science and Reading. A similar pattern is observed in the median scores. In mathematics, the right tail extends farther than the other two subjects, indicating the occurrence of maximum PV value in this subject. In reading, the box plot shows a relatively higher number of outliers at the lower range, which indicates this subject has the weakest performance among all subjects.

These distributions could be further explored in the context of school, gender, and socioeconomic background to determine if these factors influence the performance distribution.

### Relationship between Performance in Mathematics, Reading and Science with Gender

```{r}
#| code-fold: true
P4 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= Gender, y= MATH)) +
  geom_violin(color = "#459395", size = 0.6, fill= "#459395", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +
  scale_color_manual(values=c("#999999", "#E69F00")) +
  theme_minimal() +
  labs(title="Mathematics") +  
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 
  
P5 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= Gender, y= READ)) +
  geom_violin(color = "#EB7C69", size = 0.6, fill= "#EB7C69", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Reading") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 
  
P6 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= Gender, y= SCIENCE)) +
  geom_violin(color = "#FDA638", size = 0.6, fill= "#FDA638", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Science") + 
  theme(axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 

(P4 + P5 + P6) +
    plot_annotation(title= "Gender-Based Performance Comparison ",
                    theme = theme(plot.title=element_text(size= 15, hjust= 0.5)))


```

This EDA visualization utilizes violin plots with integrated box plots to compare academic performance by gender across Mathematics, Reading, and Science among Singaporean students. The violin plots suggest a roughly normal distribution for scores in all subjects, with slight variations. The 'body' of each violin is bulged in the middle, indicating a concentration of scores around the median. From the box plot, it is observed that in all subjects, the median and mean are close to each other, with the mean slightly lower, suggesting a modest left skew in the data.

In mathematics, the plots indicate males performed slightly better than females, with a slightly higher median and mean. The outliers at the higher end of the scores suggest that some students achieve excellent performance. In Reading, females markedly outperform males, evidenced by higher median and mean values, and a wider distribution at the higher end of scores. Science shows a near-identical performance between genders, with both median and mean scores closely aligned, suggesting minimal gender variation in this subject.

In general, performance across different subjects for male and females exhibits a slight variation.

### Relationship between Performance in Mathematics, Reading and Science with School Type

```{r}
#| code-fold: true
P7 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= TypeofSchool, y= MATH)) +
  geom_violin(color = "#459395", size = 0.6, fill= "#459395", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +
  scale_color_manual(values=c("#999999", "#E69F00")) +
  theme_minimal() +
  labs(title="Mathematics") +  
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 
  
P8 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= TypeofSchool, y= READ)) +
  geom_violin(color = "#EB7C69", size = 0.6, fill= "#EB7C69", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Reading") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 
  
P9 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= TypeofSchool, y= SCIENCE)) +
  geom_violin(color = "#FDA638", size = 0.6, fill= "#FDA638", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Science") + 
  theme(axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 

(P7 + P8 + P9) +
    plot_annotation(title= "School-Based Performance Comparison ",
                    theme = theme(plot.title=element_text(size= 15, hjust= 0.5)))

```

The EDA visualization comparing student performance across Mathematics, Reading, and Science reveals distinct patterns. Public schools exhibit a broader distribution of scores, suggesting a diverse range of performance levels. Conversely, private schools show a narrower spread, as evidenced by the shorter violin plots, yet with a wider width at the median, indicating a dense clustering of scores around the central tendency. This is further supported by the narrower interquartile range observed in private schools, pointing to a more consistent performance among their students.

Additionally, the mean and median scores in private schools are marginally higher than those in public schools, suggesting a slight academic advantage. Moreover, public schools display a greater number of outliers, especially in the lower score range, highlighting a segment of students with performance significantly below the average.

These insights suggest that while student performance in private schools is more consistent with a tendency towards higher achievement, public schools demonstrate a wider variability in student outcomes.

### Relationship between Performance in Mathematics, Reading and Science with Socioeconomic Status

```{r}
#| code-fold: true
P10 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= binned_ESCS, y= MATH)) +
  geom_violin(color = "#459395", size = 0.6, fill= "#459395", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +
  scale_color_manual(values=c("#999999", "#E69F00")) +
  theme_minimal() +
  labs(title="Mathematics") +  
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 8),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_discrete(breaks = unique(stu_qqq_SG_converted$binned_ESCS), 
                            labels = str_wrap(unique(stu_qqq_SG_converted$binned_ESCS),
                                              width = 10))
  
P11 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= binned_ESCS, y= READ)) +
  geom_violin(color = "#EB7C69", size = 0.6, fill= "#EB7C69", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Reading") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 8),
        axis.text.x = element_text(angle = 45, hjust = 1)) +  
  scale_x_discrete(breaks = unique(stu_qqq_SG_converted$binned_ESCS), 
                            labels = str_wrap(unique(stu_qqq_SG_converted$binned_ESCS),
                                              width = 10))
  
P12 <- ggplot(data= stu_qqq_SG_converted,
       aes(x= binned_ESCS, y= SCIENCE)) +
  geom_violin(color = "#FDA638", size = 0.6, fill= "#FDA638", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Science") + 
  theme(axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 8),
        axis.text.x = element_text(angle = 45, hjust = 1)) +  
  scale_x_discrete(breaks = unique(stu_qqq_SG_converted$binned_ESCS), 
                            labels = str_wrap(unique(stu_qqq_SG_converted$binned_ESCS),
                                              width = 10))

(P10 + P11 + P12) +
    plot_annotation(title= "Socioeconomic-Based Performance Comparison ",
                    theme = theme(plot.title=element_text(size= 15, hjust= 0.5)))

```

The visualization assesses student performance in Mathematics, Reading, and Science against socioeconomic status, categorized as Disadvantaged, Slightly Disadvantaged, Slightly Advantaged, and Advantaged. Across all subjects, a clear pattern is observable, where students from more advantaged backgrounds tend to achieve higher scores. This is evidenced by the upward shift in both the median (the horizontal line within the box) and the mean (the red dot) as we move from Disadvantaged to Advantaged.

For all three subjects, the spread of scores (indicated by the width of the violin plots) is broader for Disadvantaged and Slightly Disadvantaged groups, suggesting more variability in performance. It is also observed that the violin plots widen around the 50th percentile range, alongside a narrower interquartile range, indicating a denser concentration of scores near the median as we transition from disadvantaged to advantaged socioeconomic groups.

This visualization suggests that socioeconomic status has a consistent correlation with academic performance across different subjects. Further analysis could delve into the specific aspects of socioeconomic advantage that contribute to these patterns.

### Relationship between performances in Mathematics and school type across different socioeconomic status

```{r}
#| code-fold: true
ggplot(data= stu_qqq_SG_converted,
       aes(x=  TypeofSchool, y= MATH)) +
  geom_boxplot(width= 0.5, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  facet_grid(~ binned_ESCS) +
  labs(title= str_wrap("Comparative Analysis of Mathematics Performance by 
                       School Type Across Socioeconomic Tiers"),
       x = "Type of School") +
  theme(plot.title=element_text(size= 12, hjust= .5),
        axis.text = element_text(size= 10)) +
  theme_bw()
```

Extending the findings from the previous section, where we noted that private schools typically demonstrate higher and more consistent student performance, and students from advantaged backgrounds achieve higher scores, we now examine the impact of school type on performance across different socioeconomic levels.

The visualization offers a comparative analysis of Mathematics performance between public and private schools across different levels of socioeconomic status. For students classified as disadvantaged, private schools show a higher mean score, indicating a performance advantage in this group. As we move to the slightly disadvantaged socioeconomic status bracket, the mean scores for private schools remain higher, though the difference is less obvious.

Interestingly, a shift occurs among the slightly advantaged group, where the mean scores for public schools marginally surpass those of private schools. This trend amplifies within the advantaged category, where students in public schools outperform those in private schools, as indicated by higher mean scores. These insights suggest that school type affects educational outcomes differently across socioeconomic groups.
