---
title: "In-class Exercise 1: Now You See It!"
date: January 13, 2024
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  toc: true
editor: visual
---

## Loading R packages

In this hands-on exercise, two R packages will be used. They are:

-   [tidyverse](https://www.tidyverse.org/), and
-   [haven](https://haven.tidyverse.org/)

The code chunk used is as follows:

```{r}
pacman::p_load(tidyverse, haven)
```

## Importing PISA data

The code chunk below uses [read_sas()](https://haven.tidyverse.org/reference/read_sas.html) of [haven](https://haven.tidyverse.org/) to import PISA data into R environment.

Step 1: Use read_sas() to read SAS file

```{r}
#| eval: false
stu_qqq <- read_sas("data/cy08msp_stu_qqq.sas7bdat")
```

Step 2: Filter required data using filter()

```{r}
#| eval: false
stu_qqq_SG <- stu_qqq %>%
  filter(CNT == "SGP")
```

Step 3: Save filtered data in R Data Format
```{r}
#| eval: false
write_rds(stu_qqq_SG,
          "data/stu_qqq_SG.rds")
```

Step 4: Read saved R Data format file to ensure that it is working
```{r}
stu_qqq_SG <- read_rds("data/stu_qqq_SG.rds")
```
