---
title: "Hands-on Exercise 1"
date: January 1, 2024
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  toc: true
---

# Getting started

::: panel-tabset
## Loading R package

First, load R packages using the below code:

```{r}
pacman::p_load(tidyverse)
```

## Importing Data

The code chunk below imports exam_data.csv into R environment by using read_csv() function of readr package.

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

## Understanding data

```{r}
head(exam_data)
```

```{r}
summary(exam_data)
```
:::

# Plotting basic plot

::: panel-tabset
## Bar plot

```{r}
ggplot(data=exam_data,
       aes(x= RACE)) +
  geom_bar() 
```

## Histogram

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_histogram()    
```

## Boxplot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot() 
```
:::

# Modifying the plot

::: panel-tabset
## Bar chart

Modifying bar chart by\
(i) changing geom()\
(ii) adding labels and title\
(iii) changing the theme

```{r}
ggplot(data=exam_data,
       aes(x= RACE)) +
  geom_bar(alpha= 0.5) +
  labs(x = "Race",
       y = "Number of Students",
       title = "Race Distribution of Exam Data")+
  theme_classic()+
  theme(plot.title = element_text(hjust=0.5))
```

::: callout-note
Opacity of the bar chart has changed.\
X-axis label, y-axis label and title was added.\
A classic-looking theme, with x and y axis lines and no gridlines.
:::

## Histogram 1

Modifying histogram by changing geom()

```{r}
ggplot(data=exam_data,
       aes(x= ENGLISH)) +
  geom_histogram(bins=20,
                 color="#905C29",
                 fill="#E1C2A3",
                 size=0.8)
```

::: callout-note
Bin, color and fill is now added to the plot
:::

## Histogram 2

Modifying histogram by changing aes()

```{r}
ggplot(data=exam_data, 
       aes(x= ENGLISH, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="Grey30")
```

::: callout-note
Instead of showing count as a whole, now they are split into Male and Female
:::

## Box plot 1

Working with stat - the stat_summary() method

```{r}

ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="blue",        
               size=5)        
```

## Box plot 2

Working with stat - the geom() method

```{r}

ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",        
             fun.y="mean",           
             colour ="green",          
             size=5)        
```
:::

# More plots...

## Working with facet_wrap()

```{r}
ggplot(data=exam_data, 
       aes(x= SCIENCE)) +
  geom_histogram(bins=20) +
    facet_wrap(~ GENDER)+
  theme_bw()
```


